Sexual selection improves population fitness: a meta-analysis

### 2018-12-07 Response_to_Nature_Comms.Rmd

---
title: 'Sexual selection improves population fitness: a meta-analysis'
output:
  pdf_document: default
  word_document: default
subtitle: Response to Decision Letter
bibliography: Manuscripts/Bibliography/Bibliography.bib
---

```{r, warning=FALSE, message=FALSE, results='hide', echo = FALSE}
library(pander)
library(tidyr)
library(compute.es)
library(metafor)
library(plyr)
library(dplyr)
library(lme4)
library(car)
library(forestplot)
library(ggplot2)
library(ggthemes)
library(kableExtra)
library(ggrepel)
library(reshape2)
library(RColorBrewer)
library(ggridges)
library(rstan) #Note that installation requires some effort: dependency for brms
library(brms) 
library(backports) #seems to be a dependency
library(bayesplot)
#devtools::install_github("mvuorre/brmstools")
library(brmstools)
library(metaAidR) # install.packages("metaAidR")
library(cowplot)
library(gridExtra)
library(janitor)
#setwd("~/Documents/Uni/metaanalysis/SexualSelection")
source("I2_function.R") #Adapted function for obtaining I2 with CIs
source("Vdodge_function.R") # nice function for ggplot
source("Tidy_functions_for_brms.R") #Tidy functions for making model tables for brms

@@

>*The manuscript presents results of a meta-analysis assessing the effect of sexual selection on population fitness. This topic has been increasingly studied over the past decade or so, and the time seems ripe for a synthesis. The authors have done good job searching the literature – it includes all relevant papers I could remember, and the results should be of interest to a broad range of evolutionary biologists. I cannot expertly judge on state-of-the-art methodology, but I do have several general concerns about how the analyses were performed.*

Many thanks for your time and this valuable feedback.

>*Firstly, it seems that several male traits like to be under sexual selection, but not related to population fitness were included. I don’t think this is correct given the question being asked (i.e. whether sexual selection increases population fitness). I think traits which are directly under sexual selection (eg. attractiveness, reproductive success) should not be pooled with traits which may respond to manipulation of sexual selection indirectly (eg. male development time, survival) and can affect population fitness. Distinguishing between both types of trait could actually be revealing – e.g. could expose trade-offs between sexually selected traits and fitness components unrelated to reproductive competition (see eg. Radwan et al. 2015 Evol Biol), a thus potentially explain lower effect of sexual selection on (pooled) male traits than on female traits.*

The reviewer is correct that several traits are likely under trade-offs with other traits as sexual selection not only affects mean fitness but alters life history traits (*i.e.* sexual selection may increase male body size or attractiveness, but decrease development rate). In essence this is why our meta-analysis sought to investigate the *net* effect of sexual selection on population fitness across many traits, rather than one that may be traded off in a given direction. Importantly, to account for the variable and uncertain relationships certain traits have with fitness (as raised by the reviewer). We provide a table (Table S1) on how each trait is measured and their relationship with fitness. Given that this information was in the Supplementary Material we have **revised the manuscript to clarify the distinctions more thoroughly... Probably need to check this again**.

Additionally, the reviewers comment that male attractiveness may not be related to population fitness is correct. Because male attractiveness may be positively or negatively correlated with average fitness (although in many condition-dependent traits a positive association is assumed) we have re-classified *Male Attractiveness* as ambiguously related to fitness, rather than indirect. All subsequent models, statistics and figures now take into account this re-classification. However, given the small sample size for male attractiveness effect sizes (*n* = 6), we find only marginal changes in the results and no key changes in the key findings of the study.

@@

>*Secondly, looking at Figure 1, one notices apparent anomalies, for example significantly positive slope for male attractiveness, based on studies with average effect size close to zero. I think this (and few others) surprising estimates may result from fitting random terms across all trait types (fitted as a fixed factor, second model); I guess fitting interaction (random slopes) would not be feasible for some categories including few data points, but some of them could easily be pooled in wider categories.*

The reviewer raises questions of the analysis that produce seemingly suprising results. We agree that some of *significant* predictions in Figure 1 (whereby 95 % CIs do not overlap zero) appear surprising based on the distribution on effect sizes. In regards to the trait "male attractiveness", the 95 % CIs do not overlap zero. This is the case even though the higher weighted studies are very close to zero. However, this seeming anomoly can be explained by various reasons.

Firstly, although the effect size prediction for the trait has 95 % CIs not overlapping zero, the overall magnitude of the effect size is small (see Table exerpt), and the Bayes Factor (an estimate of the likelihood that the effect size is actually positive) is also small. The full table for all predicted effect sizes used in Figure S1 can be seen in **S8**. Below we provide the exerpt of this table and forest plot for "Male attractiveness".

Secondly, the effect size predictions were made using a single model that incorperated all 459 effect sizes, thus model predictions for each trait have narrower confidence intervals as the fitted values do not incorperate measurement error. In regards to our R code and analysis, this means that the predictions that appeared in Figure 1 are sourced from using ``fitted.brmsfit`` as opposed to ``predict.brmsfit``. In order to clarify this seeming anomaly in model predictions, some meta-analyses investigating the effect of a variable on more than one outcome model each outcome (in this case the *fitness component*) separately. Arguably, this method provides more variable model estimates as models will always have lower power due to reduced sample sizes and the incorporation of measurement error in the predictions. We have added a table to the Supplementary Material where models are independently conducted on all fitness components where n > 3 (**Table S9**). This may be of particular value for researchers using our meta-analysis as a research tool, whereby they are interested in the effect sizes from only a single trait (as opposed to our focus on population fitness as a whole).

Given, the confusion with regards to Figure 1 as it originally appeared within the manuscript we have altered the figure to a generalized forest plot by grouping the effect sizes from each of the fitness classes (ambiguous, indirect and direct). The previous Figure 1 has now been moved to the Supplementary Material with corresponding tables.

@@

```{r}
left_join(fitness.component.predictions, predictions.outcomes, by = "Fitness Component")  %>% filter(`Fitness Component` == "Male Attractiveness") %>% pander(split.table = Inf, digits = 2)
```

```{r, fig.height=2.5, fig.width=6}
pd <- position_dodgev(0.6)

@@

p.meta.male.attract <- full_dataset %>% 
  mutate(Sex = replace(as.character(Sex), Sex == "B", "Both"),
         Sex = replace(Sex, Sex == "M", "Male"),
         Sex = replace(Sex, Sex == "F", "Female"),
         Outcome.Class = factor(Outcome.Class, levels = c("Ambiguous", "Indirect", "Direct"))) %>%
  filter(Outcome == "Male Attractiveness") %>%

  ggplot(aes(x=reorder(AuthorYear, -g), y = g)) +
  scale_color_manual(values = c("Ambiguous" = "#a50f15", "Indirect" = "#fe9929", "Direct" = "#4daf4a"), 
                     name = "Relationship\nto fitness")+
  scale_shape_manual(values=c(21,22,24))+
  scale_fill_manual(values = c("Ambiguous" = "#a50f15", "Indirect" = "#fe9929", "Direct" = "#4daf4a"), 
                    name = "Relationship\nto fitness")+
  geom_linerange(aes(ymin = lowerci, 
                   ymax = upperci,
                   x = reorder(AuthorYear, -g),
                     color = Outcome.Class,
                     group = "identity"),
               show.legend = FALSE,
               position = position_dodge2(width = 0.5, preserve = "single")) +

  geom_point(aes(shape = Sex,
                 fill = Outcome.Class),
             size = 1.75, 
             color = "grey20",
             position = position_dodge2(width = 0.5),
             show.legend = F) +

@@

  #Add theme specifying text size, margins, lines etc.
  theme_bw()+

  theme(strip.text.y = element_text(angle = 0, size = 12, margin = margin(t=15, r=15, b=15, l=15)), 
        strip.background = element_rect(colour = NULL,
                                        linetype = "blank",
                                        fill = "gray90"),
        text = element_text(size=13),
        panel.spacing = unit(0.5, "lines"),
        panel.border= element_blank(),
        axis.line=element_line(), 
        panel.grid.major.x = element_line(linetype = "solid", colour = "gray95"),
        panel.grid.major.y = element_line(linetype = "solid", color = "gray95"),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(), 
        legend.text = element_text(size=12), 
        legend.title=element_text(size=12, 
                                  face = "bold"),
        axis.title.x = element_text(hjust = 0.5, size = 12))

p.meta.male.attract
```

When the effect sizes (n = 6) for male attractiveness are viewed isolated from others, one can see that the large effect size (Nelson 2013) shifts the overall predicted effect size to be marginally positive.

@@

>*Thirdly, type of trait measured explained 35% variance, but the authors do not explore this any further. However, examination of Fig. 1 suggest that some indirect/ambiguous fitness measures account for much of this heterogeneity and they generally have higher average effect sizes that direct measures (except for immunity). I’d like to see if the authors recover their main result if they only direct measures.*

**Add more in discussion about the variability amongst traits and why some might be subject to trade offs (e.g. immunity): We do kind of talk about this and an earlier draft had some of Jaceks work with regards to sexual conflict maintaing genetic variance through trade-offs. Should we add this back in?**

@@

The reviewer provides an accurate shortcoming of discussion on the observed heterogeneity between fitness components. The sixth paragraph of our discussion is raises some potential reasons why heterogeneity may be high within our meta-analysis. Additionally, it may please the reviewer to find an addition to our Supplementary Material (Table S9), now conducts independent meta-analyses and obtains $I^2$ for each of these traits. Even when metanalyses are conducted on a single trait, heterogeneity ($I^2$) is still high > 90 %.

In regards to conducting a meta-analysis on just direct fitness components, the reviewer provides a valid question of our use of direct **and** indirect fitness components for the meta-analysis investigating the effects of sex and stress. Based on the question we asked we were interested in the effect of sexual selection on many components of fitness.

@@

Nevertheless, we conducted our analysis using just direct fitness measures. We find the same key associations; that sexual selection generally elevates fitness, especially for females, with the effect magnified in stressful environments. Below we present the findings. 
**should we put this in the supp?**

Additionally, in order to further distinguish the effects of sexual selection on ambiguous, indirect and direct fitness components, we have structured the second paragraph of the results to more consistently reflect our classification system.

```{r, warning=FALSE, message=FALSE, error = FALSE, fig.align='center', fig.width = 10, fig.height=6}
model.direct.only <- rma.mv(g, V = var.g, 
                          mods = ~ 1 + Sex * Environment, 
                          random = list(~ 1 | Study.ID, 
                                       ~ 1 | Outcome,
                                       ~ 1 | Taxon), 
                          method = "REML", 
                          data = strict_dataset %>% filter(Outcome.Class == "Direct"))

@@

>*In the discussion the authors say they included the number of experimental evolution generations, but I could not find this information in methods.*

It is correct that we did not use the number of generations in our main analysis. The manuscript (Methods section) states that the number of generations. In the supplementary material we investigate the role of number of generations (the lineage is evolved for) has on fitness in the section *Meta-analyses including many moderator variables* as well as in Figure S12 and S13. Generations (specifically, log-Generations) had minimal effect on the effect sizes and was removed for the models that appeared in the main text.

>*Fig. S1 is not referred to in the main texts, is it different from Fig. 2, except that the latter contains predicted average values for fitness components?*

**No difference... kind of there for documentation on how we created it (transparency). We now get rid of the forest plot in the manuscript. However our supp info has a few plots that are in the manuscript. I think it is good to show the code we used to create it... for reproducibility**

@@

>*As a response, they come up with log response ratio (lnRR) - see*

>Hedges, L. V., J. Gurevitch, and P. S. Curtis. 1999. The meta-analysis of response ratios in experimental ecology. \textit{Ecology} 80:1150-1156.

>*I recommend that the authors use lnRR for their effect size for mean comparison as well as Hedge's $g$ to see the robustness of their conclusions for the mean.*

Reviewer 2 has rightly pointed out that the use of Hedge's *g* has been critisized for its ability to account for heterogeneity between the control and treatment groups [Osenberg_1997]. We extended our meta-analysis by including the log response ratio (*lnRR*) as well as a modified version of Hedge's g that attempts to account for this heterogeneity, this modified statistic [@bonett2009meta] can be derived from the ``metafor`` package for meta-analysis [@RN212], where it is named SMDH (standardized mean difference with heteroscedastic population variances in the two groups). There are several key findings from the use of lnRR and SMDH:

*Firstly*, Hedge's g and SMDH are highly correlated (in means and variance). This suggests that unequal variance between the treatment and control groups (heteroscedasticity) is unlikely to impact the effect size of Hedges' g.

```{r, warning=FALSE, message = FALSE, error=FALSE}
#This function writes in the correlation value
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(-2.5, 2.5, -2.5, 2.5))
  r <- cor(x, y, use="pairwise.complete.obs")
  txt <- format(c(r, 0.123456789), digits=digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0, 0, txt, cex = cex.cor * 2*r)
}

@@

es.correlation.df <- lnRR.data %>% left_join(SMD.data %>% rename(`Hedge's g` = SMD)) %>% left_join(SMDH.data)

pairs(es.correlation.df %>% select(`Hedge's g`, lnRR, SMDH), lower.panel=panel.smooth, upper.panel=panel.cor)
# pairs(es.correlation.df %>% select(SMD_var, SMDH_var), lower.panel=panel.smooth, upper.panel=panel.cor)
```

*Secondly*, The calculation of lnRR and SMDH can only be calculated for effect sizes where the study reported means, standard deviations and sample sizes (352 out of 459). This is a limitation as in our analysis we investigate the effects of sex and stress on a further restricted dataset (direct and indirect fitness traits only). Specifically, fewer effect sizes from stressful conditions are able to be calculated; making Hedges' g an advantageous effect size in comparison to lnRR.

```{r, warning=FALSE, message = FALSE, error=FALSE}
predictions.complete2 %>% select(Sex, Environment, `Hedge's g sample size` = n) %>% 
  left_join(predictions.lnRR %>% select(Sex, Environment, `lnRR sample size` = n), by = c("Sex", "Environment")) %>%
  adorn_totals("row") %>% pander()
```

*Thirdly*, if we conduct a meta-analysis using lnRR as the effect size we obtain very similar results with all major conclusions unchanged, keeping in mind that lnRR and Hedge's are on different scales; (further model results can be found within the Supplementary Material):  
+ The effect of sexual selection on population fitness is generally positive.
+ This effect is magnified for females in stressful environments with a significantly positive interaction term (stress*females).

```{r, warning=FALSE, message=FALSE, error = FALSE, fig.align='center', fig.width = 10, fig.height=6}
grid.arrange(both.plots + 
               guides(fill = FALSE) + 
               ggtitle("Comparison between Hedge's g \nand lnRR effect sizes \n"), 
             lnRR.plots + ggtitle("\n\n") + xlim(-0.5,0.75), nrow = 1, widths = c(3,4))
```

*Fourthly*, Hedges' g appears more normally distributed than lnRR, which to some extent may come as a surprise as *lnRR* should usually be more normally distributed around zero. However, given that our meta-analysis combines multiple traits, with different scales of magnitude lnRR may lead to some erroneous and biased effect sizes. Specifically, @Lajeunesse_2015 showed that *lnRR* is problematic when quantifying the outcome of studies with small sample sizes, and can yield unsuitable variance estimates when the parameter scale of the study are close to zero. The more normal distribution of the Hedges' *g* is anticipated to increase model fit.

@@

ggplot(lnRR.data, aes(sample = lnRR)) + 
  stat_qq() + 
  stat_qq_line() +
  labs(x = 'theoretical quantiles', y = "lnRR sample"),

ggplot(data = full_dataset, aes(x = g))+
  geom_density()+
  xlab("Hedges' g"),

ggplot(full_dataset, aes(sample = g)) + 
  stat_qq() + 
  stat_qq_line() +
  labs(x = 'theoretical quantiles', y = "Hedges' g sample"))
```

The lnRR is a commonly used effect size when the trait of interest has a natural zero point. Given that we did not extract the data with this in mind, it may be a further limitation of *lnRR*.

For these reasons we have decided to use Hedges' *g* as the main effect size throughout the manuscript but reference and provide equivelant statistical analysis for *lnRR* in the supplementary material.

@@

>*2 --- The authors may consider also doing another set of meta-analyses using lnVR (log variability ratio - proposed in Nakagawa et al. 2015).*

>Nakagawa, S., R. Poulin, K. Mengersen, K. Reinhold, L. Engqvist, M. Lagisz, and A. M. Senior. 2015. Meta-analysis of variation: ecological and evolutionary applications and beyond. \textit{Methods in Ecology and Evolution} 6:143-152.

>*As one can see in Figure 2, the mean results (Hedges' $g$) are a mirror image of the variance results (this makes sense CV controls for means). I would like to see what the absolute change in variances. Probably the authors can put the analysis using lnVR in the supplement. The results of this analysis can be discussed. Also, the mean-variance relationship between mean and variance (sd) should be verified (e.g. plot log mean and log sd or log variance).*

The reviewer suggestions have been included in our meta-analysis as Supplementary Material in order to maintain transparency in our conclusions. Several key points arise from the meta-analysis of variation:

*Firstly*, as the reviewer suggested we justify the use of *lnCVR* my assessing the mean-variance relationship. The major limitation of lnVR is that it does not account for the mean-variance relationship and that is why lnCVR was suggested as the favourable correction [@RN177]. Given this finding the use of *lnCVR* is preferenced over *lnVR*, as *lnVR* does not account for the mean-variance relationship.

```{r, warning=FALSE, message=FALSE, error = FALSE, fig.align='center', fig.width = 8, fig.height=3.75}
grid.arrange(full_dataset %>% ggplot(aes(x = mean.low, y = sd.low))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()+
  xlab("Mean of Control Group")+
  ylab("Standard Deviation of Control Group") +
  geom_smooth(method = "lm", colour = "red"),

@@

  full_dataset %>% ggplot(aes(x = mean.high, y = sd.high))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()+
  xlab("Mean of Treatment Group")+
  ylab("Standard Deviation of Treatment Group") +
  geom_smooth(method = "lm", colour = "red"), nrow = 1)
```

### 2018-12-07 Response_to_Nature_Comms.Rmd

*Secondly*, We conducted both Bayesian and REML meta-analyses for *lnVR*, which we present in the supplementary material and briefly discuss in the manuscript. We found a similar trend for the meta-analysis of variance using *lnCVR* and the meta-analysis of variance using *lnVR*, These are:

+  In stressful environments sexual selection leads to a decrease in the variance for females, relative to benign environments.  
+  Sexual selection reduces variance in females compared to males.

However, when using *lnVR*, the variability of fitness traits for males under sexual selection increased, *lnVR* for females in stressful environments was not statistically different from zero, and sexual selection in benign environments increased phenotypic variance for females. Model results and predictions (from the REML model) can be cound within the Supplementary Material.

```{r, warning=FALSE, message=FALSE, error = FALSE, fig.align='center', fig.width = 10, fig.height=6}
grid.arrange(both.var.plots + 
               guides(fill = FALSE) + 
               ggtitle("Comparison between lnCVR \nand lnVR effect sizes \n"), 
             both.lnVR.plots + ggtitle("\n\n"), nrow = 1, widths = c(3,4))
```

>*3 --- $I^2$ needs to be explained. $I^2$ is proposed originally here:*

>Higgins, J., and S. Thompson. 2002. Quantifying heterogeneity in a meta-analysis. \textit{Statistics in Medicine} 21:1539 - 1558.

>*But later expanded in here for mixed models (hierarchical models)*

>Nakagawa, S., and E. S. A. Santos. 2012. Methodological issues and advances in biological meta-analysis. \textit{Evolutionary Ecology} 26:1253-1274.

>*This is what the authors use. Also, it will be good to put the degree in I^2 in the context. To do this see this paper:*

>Senior, A. M., C. E. Grueber, T. Kamiya, M. Lagisz, K. O'Dwyer, E. S. A. Santos, and S. Nakagawa. 2016. Heterogeneity in ecological and evolutionary meta-analyses: its magnitude and implications. \textit{Ecology} 97:3293-3299.

We value the reviewers feedback and have incorperated all the suggestions within a paragraph within the methods of our manuscript.

>*4 --- Publication tests have been conducted on, I think, "meta-analytic" residuals results as suggested in Nakagawa and Santos 2012. One could use such residuals to conduct the trim and fill method and see how much the mean could move (see Nakagawa and Santos 2012). One should remember the funnel asymmetry could be caused by the presece of heterogeniety. See:*

>Egger, M., G. Smith, M. Schneider, and C. Minder. 1997. Bias in metaanalysis detected by a simple, graphical test. \textit{Br Med J} 315:629 - 634.

**No in-buit functionality for trim-and-fill for rma.mv in metafor... could figure out a way to do it I suppose, not sure if worth it**

>*5 --- Figure 3 - are these grey envelopes 95% CI?*

Yes they are, we have now included this information in the figure caption. Thank you for highlighting this.

>*6 --- the title - do the authors include "a systematic review" - so "a systematic reveiw and meta-analysis" - these two things are diferent - see:*

>Nakagawa, S., and R. Poulin. 2012. Meta-analytic insights into evolutionary ecology: an introduction and synthesis. \textit{Evolutionary Ecology} 26:1085-1099.

Our study does not involve a systematic review. The scope of a systematic review was larger than this meta-analysis entailed. However, we did use a systematic literature search strategy and PRISMA framework of inclusion/exclusion. **I think thewre is some unsureness here on what counts as a systematic review... I remember in medicine a meta-analysis MUST be part of a systematic review, but I guess ecology and evolution is more flexible**.

>*Hope my comments are useful.*

### 2018-12-07 Manuscripts/Manuscript_MetaAnalysis.Rmd
Response to reviewers comments

The effect size estimates we recovered were highly heterogeneous ($I^2$ = 94.7%, 95% CIs: 92.4-96.4%), reflecting the large differences in experimental procedures, study species, and fitness components included in our meta-analysis [@RN208]. Most of the observed heterogeneity stemmed from between-study differences ($I^2_{study}$ = 57.4%, 95% CIs: 38.1-76.5%), with variation between fitness components accounting for comparatively less heterogeneity ($I^2_{fitness\:components}$ = 37.4%, 95% CIs: 16.9-57.6%). The effect of taxon varied between REML and Bayesian models: the REML model suggested no effect ($I^2_{taxon}$ = 0), while the Bayesian model **suggested a probable non-zero effect** ($\sigma^2$ = 0.13 [0-0.46]). Variation among taxa is explored further in the ESM (\href{https://justincally.github.io/SexualSelection/#restricted_analysis_to_just_the_higher-quality_data}{Figure S2, Table S11}).  \phantomsection

@@

As in the meta-analysis of trait means, there was high heterogeneity in the estimates of *lnCVR* ($I^2$ = 95.8%, 95% CIs: 93.4-97.5%). Heterogeneity in the dataset was due to variability between studies ($I^2_{study}$ = 34.8 %, 95% CIs: 18.8 % to 56.1 %), fitness components ($I^2_{fitness\:components}$ = 48.8%, 95% CIs: 23.3-70.1%) and taxon ($I^2_{taxon}$ = 12.2, 95% CIs: 2.07-29.8%).  \phantomsection

@@

**Additionally, using studies that presented means, standard deviations and sample sizes (n = 352) we were able to calculate an alternative effect size: the log-response ratio (lnRR) CITE. The lnRR was used as a supplement to Hedges' g as it relaxes the assumption in equal variances between control and treatment groups (homodescdandsavn).** However, lnRR was not able to be calculated from studies that

@@

Thirdly, we measured the impact of environment, sex and their interaction on the effect size (Hedges’ *g*, *lnRR* and *lnCVR*) associated with the manipulation of sexual selection, by fitting these predictors as moderators in a pair of separate mixed-effects meta-analyses. These meta-analyses were restricted to effect sizes calculated from ‘unambiguous’ outcomes (*i.e.* those scored as being directly or indirectly related to population fitness), as well as those where we were able to define the environmental conditions as either stressful or benign (Hedges' *g*: *n* = 330; *lnRR* & *lnCVR*: *n* = 269). We again fit study ID, fitness component, and taxon as random/group level effects. Models investigating other moderators such as number of generations and blinding are presented in the \href{https://justincally.github.io/SexualSelection/}{ESM}.

@@

![Tests of publication bias are mixed and suggest publication bias may be present. (**a**) Inspection and statistical tests of the funnel plot reveal large amounts of heterogeneity in the dataset with asymmetry from increased low-powered, large effect studies. (**b**) No significant correlation exists between journal impact factor and effect size, (**c**) additionally there is no significant correlation between effect size and year of publication when testing for the time-lag bias. For (**b**) and (**c**) point size is proportional to the precision of the effect size (*i.e.* the inverse of its variance).\label{publication_bias}](Figures/publicationbias.pdf)

>>>>

The effect size estimates we recovered were highly heterogeneous ($I^2$ = 94.7%, 95% CIs: 92.4-96.4%), reflecting the large differences in experimental procedures, study species, and fitness components included in our meta-analysis [@RN208; @Senior_I2_2016]. Most of the observed heterogeneity stemmed from between-study differences ($I^2_{study}$ = 57.4%, 95% CIs: 38.1-76.5%), with variation between fitness components accounting for comparatively less heterogeneity ($I^2_{fitness\:components}$ = 37.4%, 95% CIs: 16.9-57.6%). The effect of taxon varied between REML and Bayesian models: the REML model suggested no effect ($I^2_{taxon}$ = 0), while the Bayesian model **suggested a probable non-zero effect** ($\sigma^2$ = 0.13 [0-0.46]). Variation among taxa is explored further in the ESM (\href{https://justincally.github.io/SexualSelection/#restricted_analysis_to_just_the_higher-quality_data}{Figure S2, Table S11}).  \phantomsection

@@

As in the meta-analysis of trait means, there was high heterogeneity in the estimates of *lnCVR* ($I^2$ = 99.2%, 95% CIs: 98.8-99.6%). Heterogeneity in the dataset was due to variability between studies ($I^2_{study}$ = 36.1 %, 95% CIs: 19.3 % to 59 %), fitness components ($I^2_{fitness\:components}$ = 50.4%, 95% CIs: 24.2-72.1%) and taxon ($I^2_{taxon}$ = 12.7, 95% CIs: 2.14-31%).  \phantomsection

@@

**Additionally, using studies that presented means, standard deviations and sample sizes (n = 352) we were able to calculate an alternative effect size: the log-response ratio (lnRR) CITE. The lnRR was used as a supplement to Hedges' g as it relaxes the assumption in equal variances between control and treatment groups (homoscedasticity).** However, lnRR was not able to be calculated from studies that

@@

Thirdly, we measured the impact of environment, sex and their interaction on the effect size **(Hedges’ *g*, *lnRR*, *lnCVR* and *lnVR* in the ESM)** associated with the manipulation of sexual selection, by fitting these predictors as moderators in a pair of separate mixed-effects meta-analyses. These meta-analyses were restricted to effect sizes calculated from ‘unambiguous’ outcomes (*i.e.* those scored as being directly or indirectly related to population fitness), as well as those where we were able to define the environmental conditions as either stressful or benign **(Hedges' *g*: *n* = 330; *lnRR*, *lnCVR* & *lnVR*: *n* = 269)**. We again fit study ID, fitness component, and taxon as random/group level effects. Models investigating other moderators such as number of generations and blinding are presented in the \href{https://justincally.github.io/SexualSelection/}{ESM}.

**For our meta-analysis investigating the effects of environment and sex on the magnitude and variance of fitness-related traits (*Hedges' g* and *lnCVR*) we provide estimates of heterogenity present in the dataset. We use the $I^2$ statistic as an estimate of the proportion of total variation in study estimates that is due to heterogeneity [@Higgins_2002]. $I^2$ is prefered over other statistics as it is independent of sample size, is easily interperatable (pecentage) and can be partitioned between random effects [@RN208]. Within ecology and evolution heterogeneity in datasets is often high, with the mean $I^2$ from 86 studies above 90 % [@Senior_I2_2016].**

@@

![Tests of publication bias are mixed and suggest publication bias may be present. (**a**) Inspection and statistical tests of the funnel plot reveal large amounts of heterogeneity in the dataset with asymmetry from increased low-powered, large effect studies. (**b**) No significant correlation exists between journal impact factor and effect size, (**c**) additionally there is no significant correlation between effect size and year of publication when testing for the time-lag bias. For (**b**) and (**c**) point size is proportional to the precision of the effect size (*i.e.* the inverse of its variance). Grey envelopes in (**b**) and (**c**) represent the 95 % CI of the linear regression.\label{publication_bias}](Figures/publicationbias.pdf)

### 2018-12-07 meta_analysis_sexual_selection_pop_fitness.Rmd
Response to reviewers comments

From this data we can generate a forest plot with the following:

@@

  theme(strip.text.y = element_text(angle = 0, size = 8, margin = margin(t=15, r=15, b=15, l=15)), 
        strip.background = element_rect(colour = NULL,
                                        linetype = "blank",
                                        fill = "gray90"),
        text = element_text(size=11),
        panel.spacing = unit(0.5, "lines"),
        panel.border= element_blank(),
        axis.line=element_line(), 
        panel.grid.major.x = element_line(linetype = "solid", colour = "gray95"),
        panel.grid.major.y = element_line(linetype = "solid", color = "gray95"),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(), 
        legend.text = element_text(size=12), 
        legend.title=element_text(size=12, 
                                  face = "bold"),
        axis.title.x = element_text(hjust = 0.5, size = 14))
```

<img src="figures/ForestPlot_size_reduced.pdf" alt=" "  width="1650" height="2100">

@@

```{r}
I2(model.complete2, strict_dataset$var.g) %>% pander(digits = 3)
```
These values indicate that 61.5 % of total heterogeneity is due to the **between study** heterogeneity and 33.5 % for **between outcome** heterogeneity between different outcomes. The total _I^2^_  is 95 %, a reasonably high _I^2^_ value. However this is relatively common in Ecology and Evolution total.

@@

Similar to the meta-analysis on Hedges' g we can obtain _I^2^_ for lnCVR REML model. In this case (compared to Hedges' g) we see Taxon has more of a variable effect on overall _I^2^_ estimates, although the overall _I^2^_ remains relatively similar (95.4 %).

```{r}
I2(variance.model, strict_dataset$var.g) %>% pander(digits = 3)
```
_________

>>>>

@@

  theme(strip.text.y = element_text(angle = 0, size = 8, margin = margin(t=15, r=15, b=15, l=15)), 
        strip.background = element_rect(colour = NULL,
                                        linetype = "blank",
                                        fill = "gray90"),
        text = element_text(size=11),
        panel.spacing = unit(0.5, "lines"),
        panel.border= element_blank(),
        axis.line=element_line(), 
        panel.grid.major.x = element_line(linetype = "solid", colour = "gray95"),
        panel.grid.major.y = element_line(linetype = "solid", color = "gray95"),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(), 
        legend.text = element_text(size=12), 
        legend.title=element_text(size=12, 
                                  face = "bold"),
        axis.title.x = element_text(hjust = 0.5, size = 14))

knitr::include_graphics(path = "figures/ForestPlot_size_reduced.png") 
```

@@

```{r}
I2(model.complete2, strict_dataset$var.g) %>% pander(digits = 3)
```
These values indicate that 57.4 % of total heterogeneity is due to the **between study** heterogeneity and 37.4 % for **between outcome** heterogeneity between different outcomes. The total _I^2^_  is 94.7 %, a reasonably high _I^2^_ value. However this is relatively common in Ecology and Evolution total [@Senior_I2_2016].

@@

Similar to the meta-analysis on Hedges' g we can obtain _I^2^_ for lnCVR REML model. In this case (compared to Hedges' g) we see Taxon has more of a variable effect on overall _I^2^_ estimates and the total _I^2^_ is larger.

```{r}
I2(variance.model, na.omit(strict_dataset$var.lnCVr)) %>% pander(digits = 3)
```
_________
